---
title: "Regutools: an R package for data extraction from RegulonDB"
author: "Jose Alquicira Hernandez, Carmina Barberena Jonas, Jesus Emiliano Sotelo Fonseca "
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Introduction

**RegulonDB** is a relational database that offers, in an organized and computable form, updated knowledge on transcriptional regulation in *Escherichia coli K-12*. **RegulonDB** has more than 70 tables currently available and as it continues to grow manual information extraction gets complex. This R package provides a simplified version of **RegulonDB**, as well as querying functions that make data exploration and extraction easy and automatic.

## Exploring the database

The simplified version of RegulonDB consists in a small number of  datasets each of them representing a regulation object. To see a list of available datasets, 

```{r echo = TRUE}
library(regutools)
ListDatasets()
```

## Columns of a table

```{r echo = TRUE}
head(ListAttributes("GENE"))
```

## Column description

```{r echo = TRUE}
head(ListAttributes("GENE",
               comments = TRUE))
```

## Retrieving data
```{r echo = TRUE}
head(GetAttr(attributes=c("posright","name"),
        dataset="GENE"))
```

## Using filters
```{r echo = TRUE}
GetAttr(attributes=c("posright","name"),
        filters=list("name"=c("araC","crp","lacI")),
        dataset="GENE")
```

## Interval filter
```{r echo = TRUE}
GetAttr(attributes=c("posright","name"),
        filters=list("posright"=c(1,5000)),
        interv=c("posright"),
        dataset="GENE")
```

## Partial match
```{r echo = TRUE}
head(GetAttr(attributes=c("posright","name"),
        filters=list("name"="ara"),
        partialmatch="name",
        dataset="GENE"))
```

## Complex queries
```{r echo = TRUE}
dim(GetAttr(attributes = c("name", "strand", "posright", "product_name"), 
           dataset = "GENE",
           filters = list(name=c("ARA"),
                          product_name=c("Ara"),
                          strand=c("forward"),
                          posright=c("2000","4000000")
           ),
           and=FALSE,
           partialmatch = c("name", "product_name") ,
           interv="posright" ))
```
## Complex queries
```{r echo = TRUE}
GetAttr(attributes = c("name", "strand", "posright", "product_name"), 
           dataset = "GENE",
           filters = list(name=c("ARA"),
                          product_name=c("Ara"),
                          strand=c("forward"),
                          posright=c("2000","4000000")
           ),
           and=TRUE,
           partialmatch = c("name", "product_name") ,
           interv="posright" )
```

## Gene regulation 
```{r echo = TRUE}
GetGeneRegulation(c("araC","fis","crp"))
```

## Table format
```{r echo = TRUE}
GetGeneRegulation(c("araC","fis","crp"),
                  format="table")
```

## Retrieve whole network
```{r echo = TRUE}
head(GetNetwork())
```

##  TF-TF network
```{r echo = TRUE}
head(GetNetwork(type = "TF-TF"))
```

## Preguntas
¿El gen "celA" existe?
```{r echo = TRUE, error = TRUE}
head(GetAttr(attributes=c("name"),
             filters=list("name"="celA"),
             dataset="GENE"))
```

```{r echo = TRUE}
head(GetAttr(attributes=c("name","synonyms"),
             filters=list("synonyms"="celA"),
             partialmatch = "synonyms",
             dataset="GENE"))
```
