---
title: 'Regutools: an R package for data extraction from RegulonDB'
author: "Jose Alquicira, Carmina Barberena, Heladia Salgado, Emiliano Sotelo"
date: "16 de mayo de 2017"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- There is no tool implemented in a **data analysis-oriented** language to make data retrieval from **RegulonDB** easy and automatic. 
- Here, we present **Regutools**, an **R package** that facilitates **data extraction**  from RegulonDB. 

## Exploring the database

```{r echo = TRUE}
library(regutools)
ListDatasets()
```

## Columns of a table

```{r echo = TRUE}
head(ListAttributes("GENE"))
```

## Column description

```{r echo = TRUE}
head(ListAttributes("GENE",
               comments = TRUE))
```

## Retrieving data
```{r echo = TRUE}
head(GetAttr(attributes=c("posright","name"),
        dataset="GENE"))
```

## Using filters
```{r echo = TRUE}
GetAttr(attributes=c("posright","name"),
        filters=list("name"=c("araC","crp","lacI")),
        dataset="GENE")
```

## Interval filter
```{r echo = TRUE}
GetAttr(attributes=c("posright","name"),
        filters=list("posright"=c(1,5000)),
        interv=c("posright"),
        dataset="GENE")
```

## Partial match
```{r echo = TRUE}
head(GetAttr(attributes=c("posright","name"),
        filters=list("name"="ara"),
        partialmatch="name",
        dataset="GENE"))
```

## Complex queries
```{r echo = TRUE}
dim(GetAttr(attributes = c("name", "strand", "posright", "product_name"), 
           dataset = "GENE",
           filters = list(name=c("ARA"),
                          product_name=c("Ara"),
                          strand=c("forward"),
                          posright=c("2000","4000000")
           ),
           and=FALSE,
           partialmatch = c("name", "product_name") ,
           interv="posright" ))
```
## Complex queries
```{r echo = TRUE}
GetAttr(attributes = c("name", "strand", "posright", "product_name"), 
           dataset = "GENE",
           filters = list(name=c("ARA"),
                          product_name=c("Ara"),
                          strand=c("forward"),
                          posright=c("2000","4000000")
           ),
           and=TRUE,
           partialmatch = c("name", "product_name") ,
           interv="posright" )
```

## Gene regulation 
```{r echo = TRUE}
GetGeneRegulation(c("araC","fis","crp"))
```

## Table format
```{r echo = TRUE}
GetGeneRegulation(c("araC","fis","crp"),
                  format="table")
```

## Retrieve whole network
```{r echo = TRUE}
head(GetNetwork())
```

##  TF-TF network
```{r echo = TRUE}
head(GetNetwork(type = "TF-TF"))
```

## Preguntas
¿El gen "celA" existe?
```{r echo = TRUE, error = TRUE}
head(GetAttr(attributes=c("name"),
             filters=list("name"="celA"),
             dataset="GENE"))
```

```{r echo = TRUE}
head(GetAttr(attributes=c("name","synonyms"),
             filters=list("synonyms"="celA"),
             partialmatch = "synonyms",
             dataset="GENE"))
```

## ¿Cuantas TU's hay en un solo operon? 

```{r echo=TRUE}
hist(GetAttr(attributes="total_tu",dataset="OPERON")[,1],
     main = NA,
     xlab = "TU's in a single operon")
```

## ¿Cuantas TU's hay en un solo operon? 

```{r echo=TRUE}
tu<-GetAttr(attributes="total_tu",dataset="OPERON")[,1]
tu<-tu[tu>3]
hist(tu, main = NA,xlab = "TU's in a single operon")
```

## ¿Cuantas TU's hay en un solo operon? 

```{r echo=TRUE}
GetAttr(attributes=c("name","total_promoter","sigma_name","tf_name"),
        filters = list("total_tu"=12),
        dataset="OPERON")
```

## ¿Cuántos reguladores existen por efecto?
```{r echo=TRUE}
pie(table(GetNetwork()$effect),
    col=c("green","blue","red"),
    main = "Number of regulations by effect")
```


##  Validaciones
```{r echo=TRUE , error=TRUE }
GetAttr(attributes = c("name", "stran", "posright", "product_name"), 
           dataset = "GENE",
           filters = list(name=c("ARA"),
                          product_name=c("Ara"),
                          strand=c("forward"),
                          posright=c("2000","4000000")
           ),
           and=FALSE,
           partialmatch = c("name", "product_name") ,
           interv="posright" )
```

## Perspectivas
- Bioconductor
- Artículo
- Shiny

